Synthesising Termination and Non-Termination Proofs for Bit-Vector Programs

### 2014-10-17 papers/termination/termination-esop/paper.tex
added back the proof language

%% \subsection{The Proof Language}

%% The proofs generated by the synthesiser take the form of imperative programs
%% written in the language $\mathcal{L}$.  An $\mathcal{L}$-program is a list
%% of instructions, each of which matches one of the patterns shown in
%% Figure~\ref{fig:l-language}.  An instruction has an opcode (such as
%% \verb|add| for addition) and one or more operands.  An operand is either a
%% constant, one of the program's inputs or the result of a previous
%% instruction.  The $\mathcal{L}$ language has various arithmetic and logical
%% operations, as well as basic branching in the form of the \verb|ite|
%% (if-then-else) instruction.

%% % The (non-)termination proofs are arbitrary expressions over the grammar given in Figure~\ref{fig:l-language}.  
%% % %written in a simple RISC-like language $\mathcal{L}$, whose syntax is given in Fig.~\ref{fig:l-language}.  
%% % \todo{update the figure and make it look more like a grammar.}

%% \begin{figure}
%%  \begin{center}
%%  \resizebox{\linewidth}{!}{%
%%   \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,
%%  semithick, initial text=]

%%   \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=1cm, minimum width=1.5cm},
%%           row sep=2cm, column sep=2.5cm, ampersand replacement=\&] {
%%    \node (synth) {Synthesise};
%%    \&
%%    \node (verif) {Verify}; %\\
%%    %\node[draw=none] {};
%%    \&
%%    \node[ellipse] (done) {Done}; \\
%%   };

%%    \path
%%     (synth) edge [bend left] node {Candidate program} (verif)
%%     (verif) edge [bend left] node {Counterexample input} (synth)
%%     (verif) edge node {Valid} (done);
%%  \end{tikzpicture}
%%  }
%%  \end{center}

%%  \caption{Abstract synthesis refinement loop
%%  \label{fig:abstract-refinement}}
%% \end{figure}

%% \begin{figure}
%% \begin{center}
%% {\small

%% \setlength{\tabcolsep}{14pt}
%% Integer arithmetic instructions:

%% \begin{tabular}{llll}
%%  \verb|add a b| & \verb|sub a b| & \verb|mul a b| & \verb|div a b| \\
%%  \verb|neg a| &   \verb|mod a b| & \verb|min a b| & \verb|max a b|
%% \end{tabular}

%% \medskip

%% Bitwise logical and shift instructions:

%% \begin{tabular}{lll}
%%  \verb|and  a b| & \verb|or   a b| & \verb|xor a b| \\
%%  \verb|lshr a b| & \verb|ashr a b| & \verb|not a|
%% \end{tabular}

%% \medskip

%% Unsigned and signed comparison instructions:

%% \begin{tabular}{lll}
%%  \verb|le  a b| & \verb|lt  a b| & \verb|sle  a b| \\
%%  \verb|slt a b| & \verb|eq  a b| & \verb|neq  a b| \\
%% \end{tabular}

%% \medskip

%% Miscellaneous logical instructions:

%% \begin{tabular}{lll}
%%  \verb|implies a b| & \verb|ite a b c| &  \\
%% \end{tabular}

%% \medskip

%% \setlength{\tabcolsep}{12pt}

%% Floating-point arithmetic:

%% \begin{tabular}{llll}
%%  \verb|fadd a b| & \verb|fsub a b| & \verb|fmul a b| & \verb|fdiv a b|
%% \end{tabular}

%% }
%% \end{center}

%%  \caption{The language $\mathcal{L}$}
%%  \label{fig:l-language}
%% \end{figure}

>>>>

\subsection{The Proof Language}

The proofs generated by the synthesiser take the form of imperative programs
written in the language $\mathcal{L}$.  An $\mathcal{L}$-program is a list
of instructions, each of which matches one of the patterns shown in
Figure~\ref{fig:l-language}.  An instruction has an opcode (such as
\verb|add| for addition) and one or more operands.  An operand is either a
constant, one of the program's inputs or the result of a previous
instruction.  The $\mathcal{L}$ language has various arithmetic and logical
operations, as well as basic branching in the form of the \verb|ite|
(if-then-else) instruction.

% The (non-)termination proofs are arbitrary expressions over the grammar given in Figure~\ref{fig:l-language}.  
% %written in a simple RISC-like language $\mathcal{L}$, whose syntax is given in Fig.~\ref{fig:l-language}.  
% \todo{update the figure and make it look more like a grammar.}

\begin{figure}
 \begin{center}
 \resizebox{\linewidth}{!}{%
  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,
 semithick, initial text=]

  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=1cm, minimum width=1.5cm},
          row sep=2cm, column sep=2.5cm, ampersand replacement=\&] {
   \node (synth) {Synthesise};
   \&
   \node (verif) {Verify}; %\\
   %\node[draw=none] {};
   \&
   \node[ellipse] (done) {Done}; \\
  };

   \path
    (synth) edge [bend left] node {Candidate program} (verif)
    (verif) edge [bend left] node {Counterexample input} (synth)
    (verif) edge node {Valid} (done);
 \end{tikzpicture}
 }
 \end{center}

 \caption{Abstract synthesis refinement loop
 \label{fig:abstract-refinement}}
\end{figure}

\begin{figure}
\begin{center}
{\small

\setlength{\tabcolsep}{14pt}
Integer arithmetic instructions:

\begin{tabular}{llll}
 \verb|add a b| & \verb|sub a b| & \verb|mul a b| & \verb|div a b| \\
 \verb|neg a| &   \verb|mod a b| & \verb|min a b| & \verb|max a b|
\end{tabular}

\medskip

Bitwise logical and shift instructions:

\begin{tabular}{lll}
 \verb|and  a b| & \verb|or   a b| & \verb|xor a b| \\
 \verb|lshr a b| & \verb|ashr a b| & \verb|not a|
\end{tabular}

\medskip

Unsigned and signed comparison instructions:

\begin{tabular}{lll}
 \verb|le  a b| & \verb|lt  a b| & \verb|sle  a b| \\
 \verb|slt a b| & \verb|eq  a b| & \verb|neq  a b| \\
\end{tabular}

\medskip

Miscellaneous logical instructions:

\begin{tabular}{lll}
 \verb|implies a b| & \verb|ite a b c| &  \\
\end{tabular}

\medskip

\setlength{\tabcolsep}{12pt}

Floating-point arithmetic:

\begin{tabular}{llll}
 \verb|fadd a b| & \verb|fsub a b| & \verb|fmul a b| & \verb|fdiv a b|
\end{tabular}

}
\end{center}

 \caption{The language $\mathcal{L}$}
 \label{fig:l-language}
\end{figure}
